The formal verification of the ctm approach to forcing \thanks{Supported by Secyt-UNC project 33620180100465CB and Conicet.}

### 2022-09-20 Independence_CH/replacement_instances.tex
Replacement instances in tabular format.

In \isa{instances1{\uscore}fms}:

\begin{itemize}
\item Transitive closure:
  \begin{itemize}
  \item
    \isa{eclose{\uscore}closed{\uscore}fm}.

    To prove closure under iteration of $X\mapsto\union X$.
  \item
    \isa{eclose{\uscore}abs{\uscore}fm}.

    Auxiliary instance used to show absoluteness.
  \end{itemize}
%% The instances so far
%% are needed to interpret locale
%% \locale{M{\uscore}eclose}.
\item \isa{wfrec{\uscore}rank{\uscore}fm}.

  For $\in$-rank.
  %
\item \isa{transrec{\uscore}VFrom{\uscore}fm}.

  For the cumulative hierarchy (rank initial segments).
\end{itemize}

%% The last two and next pair have the same syntactic structure, because
%% they are definitions by well-founded recursion.
In \isa{instances2{\uscore}fms} (for cardinal arithmetic):
\begin{itemize}
\item Ordertypes:
  \begin{itemize}
  \item
    \isa{wfrec{\uscore}ordertype{\uscore}fm}.

    Well-founded recursion for the construction of ordertypes.
    %
  \item
    \isa{omap{\uscore}replacement{\uscore}fm}.

    Auxiliary instance for the definition of ordertypes.
  \end{itemize}
\item Aleph:
  \begin{itemize}
    %
  \item
    \isa{ordtype{\uscore}replacement{\uscore}fm}.

    Replacement through the ordertype function, for Hartogs' Theorem.
    %
  \item
    \isa{wfrec{\uscore}Aleph{\uscore}fm}.

    The well-founded recursion to define Aleph.
  \end{itemize}
\end{itemize}

In \isa{instances{\uscore}ground{\uscore}fms}:

\begin{itemize}
\item \isa{wfrec{\uscore}Hcheck{\uscore}fm}.

  Well-founded recursion to define check-names.
  %
\item \isa{wfrec{\uscore}Hfrc{\uscore}at{\uscore}fm}.

  Well-founded recursion for the definition of forcing for atomic formulas.
  %
\item
  \isa{lam{\uscore}replacement{\uscore}check{\uscore}fm}.

  Replacement through $x\mapsto \lb x,\check{x}\rb$ (for the
  definition of $\punto{G}$).
  %
\end{itemize}

In our development, the $\isa{ground{\uscore}repl{\uscore}fm}$
function maps instances to \dots. All ground replacement instances
appear in the locale \locale{M{\uscore}ZF3} and form the set
\isa{instances3{\uscore}fms}. These are expressed using the following
predicate:
\[
  \isa{ground{\uscore}replacement{\uscore}assm}(M,\mathit{env},\phi)
\]
defined as
$\isa{replacement{\uscore}assm}(M,\mathit{env},\isa{ground{\uscore}repl{\uscore}fm}(\phi))$.

In 
\isa{instances{\uscore}ground{\uscore}notCH{\uscore}fms}:
\begin{itemize}
\item
  \isa{recursive{\uscore}construction{\uscore}fm}.

  Recursive construction of sets using a choice function.
  %
\item
  \isa{recursive{\uscore}construction{\uscore}abs{\uscore}fm}.

  Absoluteness of the previous construction.
\end{itemize}
%
\begin{itemize}
\item
  \isa{dc{\uscore}abs{\uscore}fm}.

  Absoluteness of the recursive construction in the proof of the
  Dependent Choices from $\AC$.
\end{itemize}

>>>>

In Table~\ref{tab:instances1} we show the name of the fourteen formulas
involved in the twenty two instances of replacement needed in our
mechanization. The formulas marked with ($\dagger$) are needed twice:
one by themselves and the other as the argument for $\isa{ground{\uscore}repl{\uscore}fm}$.

In our development, the $\isa{ground{\uscore}repl{\uscore}fm}$
function maps $\phi$ to $\psi$ as described in
Sect.~\ref{sec:repl-instances}. These instances form the set
\isa{instances3{\uscore}fms}.

% These are expressed using the following predicate:
% \[
%   \isa{ground{\uscore}replacement{\uscore}assm}(M,\mathit{env},\phi)
% \]
% defined as
% $\isa{replacement{\uscore}assm}(M,\mathit{env},\isa{ground{\uscore}repl{\uscore}fm}(\phi))$.
\newcommand{\groundRepl}{\ensuremath{{}^\dagger}}
\newcommand{\replInstSet}[1]{\multicolumn{3}{@{}l}{\hspace{0.3em}{$\mathbf{\mathit{#1}}$}}}

\begin{table}[!h]
\centering
\begin{threeparttable}
\begin{tabular}{r<{\stepcounter{replInstCount}\thereplInstCount.} >{\hspace{2pt}}l @{\hspace{1ex}} p{6cm}}
  \toprule
  \multicolumn{1}{r}{No.} & name & {purpose} \\
  \midrule
  \replInstSet{instances1{\uscore}fms}\\
%  \addlinespace
  & \isa{eclose{\uscore}closed{\uscore}fm} \groundRepl & closure under iteration of $X\mapsto\union X$. \\
  & \isa{eclose{\uscore}abs{\uscore}fm} \groundRepl & Absoluteness of the previous construction.\\
%  \addlinespace
  & \isa{wfrec{\uscore}rank{\uscore}fm} \groundRepl &  For $\in$-rank.\\
%  \addlinespace
  & \isa{transrec{\uscore}VFrom{\uscore}fm} \groundRepl & For rank initial segments.\\
  \midrule
  \replInstSet{instances2{\uscore}fms}\\
%  \addlinespace
  & \isa{wfrec{\uscore}ordertype{\uscore}fm} \groundRepl & Well-founded recursion for the construction of ordertypes. \\
  & \isa{omap{\uscore}replacement{\uscore}fm} \groundRepl & Auxiliary instance for the definition of ordertypes. \\
%  \addlinespace
  & \isa{ordtype{\uscore}replacement{\uscore}fm} \groundRepl& Replacement through the ordertype function, for Hartogs' Theorem.\\
%  \addlinespace
  & \isa{wfrec{\uscore}Aleph{\uscore}fm} \groundRepl& Well-founded recursion to define Aleph.\\
  \midrule
  \replInstSet{instances{\uscore}ground{\uscore}fms}\\
%  \addlinespace
  & \isa{wfrec{\uscore}Hcheck{\uscore}fm} & Well-founded recursion to define check.\\
  & \isa{wfrec{\uscore}Hfrc{\uscore}at{\uscore}fm}. & Well-founded recursion for the definition of forcing for atomic formulas.\\
  & \isa{lam{\uscore}replacement{\uscore}check{\uscore}fm} & Replacement through $x\mapsto \lb x,\check{x}\rb$ (for the
  definition of $\punto{G}$).\\
  \midrule
  \replInstSet{instances{\uscore}ground{\uscore}notCH{\uscore}fms}\\
%  \addlinespace
  &   \isa{recursive{\uscore}construction{\uscore}fm} &
  Recursive construction of sets using a choice function.\\
  & \isa{recursive{\uscore}construction{\uscore}abs{\uscore}fm} &
  Absoluteness of the previous construction \\
  \midrule
  \replInstSet{instances{\uscore}ground{\uscore}CH{\uscore}fms}\\
%  \addlinespace
  & \isa{dc{\uscore}abs{\uscore}fm} &  Absoluteness of the recursive construction in the proof of
  Dependent Choice from $\AC$. \\
  \midrule
  \replInstSet{instances3{\uscore}fms}\\
%  \addlinespace
  \multicolumn{1}{r}{15-22.} & $\isa{ground{\uscore}repl{\uscore}fm}(\phi)$ & one for each formula $\phi$ marked with $\dagger$ \\
  \bottomrule
\end{tabular}
\caption{Replacement Instances used in our mechanization}
\label{tab:instances1}
\end{threeparttable}
\end{table}
